---
title: "R_4_categoricals_stats"
author: "Mike Montes"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## 1. Load and call dataset
```{r}
library(readr)
loan <- read_csv("C:/Users/mmsax/School_Portfolio/Coding_Skills/loan.csv")
head(loan, 2)
```

## 2. Show continuous & categorical variables in the dataset.
### I made this a little fancier to practice.
### I could have switched all the characters to factors here, but I reserved not doing it yet.  I realize that integers and logicals could be included in !numeric.
```{r}
categoricals <- sapply(loan, class) != 'numeric'
cat_names <-  names(loan[categoricals])
print('The categorical variables are:')
print(cat_names)

numericals <- sapply(loan, class) == 'numeric'
num_names <- names(loan[numericals])
print('The numerical variables are:')
print(num_names)
```

## 3.  Calculate the minimum, maximum, mean, median, standard deviation and three quartiles (25th, 50th and 75th percentiles) of loan_amnt.
### I practiced old and new commands.
### I could have done the quartiles separately, but I wanted to put them together without using a loop so I had to learn 'collapse' so the 2nd and 3rd values wouldn't get cut off

### The boring method is first but is always useful
```{r}
(summary(loan$loan_amnt))

print(paste('The minimum loan_amount value is: $', (min(loan$loan_amnt))))
print(paste('The maximum loan_amount value is: $', (max(loan$loan_amnt))))
print(paste('The mean loan_amount value is: $', round(mean(loan$loan_amnt), 2)))
print(paste('The median loan_amount value is: $', (median(loan$loan_amnt))))
print(paste('The standard deviation of the loan_amount values is: $', round(sd(loan$loan_amnt), 2)))
# I could have done these separately, but I wanted to put them together without using a
# loop so I had to learn 'collapse' so the 2nd and 3rd values wouldn't get cut off
print(paste('The 25th, 50th and 75th percentiles of the loan_amount values are: $', 
            paste(quantile(loan$loan_amnt, probs = c(0.25, 0.5, 0.75)), collapse = ", ")))
```


## 4.  Calculate the minimum, maximum, mean, median, standard deviation and three quartiles (25th, 50th and 75th percentiles) of int_rate.
```{r}
(summary(loan$int_rate))

# Practicing old and new commands:
print(paste('The minimum int_rate value is: $', (min(loan$int_rate))))
print(paste('The maximum int_rate value is: $', (max(loan$int_rate))))
print(paste('The mean int_rate value is: $', round(mean(loan$int_rate), 2)))
print(paste('The median int_rate value is: $', (median(loan$int_rate))))
print(paste('The standard deviation of the int_rate values is: $', round(sd(loan$int_rate), 2)))

print(paste('The 25th, 50th and 75th percentiles of the int_rate values are: $', 
            paste(quantile(loan$int_rate, probs = c(0.25, 0.5, 0.75)), collapse = ", ")))

```

## 5.  Calculate the correlation coefficient of int_rate and installment and detemine if they have a strong relationship.
```{r}
print(paste('The correlation between int_rate and installment is:', round(cor(loan$int_rate, loan$installment), 5)))
print("This is a very low correlation value, so they do not have a strong relationship")

```

## 6.  Frequency table and mode of term.
```{r}
is.factor(loan$term)
class(loan$term)
loan$term <-  as.factor(loan$term)
is.factor(loan$term)
levels(loan$term)

print('The frequency table for loan$term:')
print(table(loan$term))

print(paste('The mode of term is:', names(sort(table(loan$term), decreasing = TRUE))[1]))
```

## 7. The proportion table and mode of loan_status.
```{r}
is.factor(loan$loan_status)
class(loan$loan_status)
loan$loan_status <-  as.factor(loan$loan_status)
is.factor(loan$loan_status)
levels(loan$loan_status)

print('The proportion table for loan$status:')
print(proportions((table((loan$loan_status)))))

  print(paste('The mode of loan_status is:', names(sort(table(loan$loan_status), decreasing = TRUE))[1]))

```

## 8. The cross table of term and loan_status and proportions by row and column respectively.
```{r}
xtabs(~term + loan_status, data = loan)

print('Proportion table by row')
prop.table((xtabs(~term + loan_status, data = loan)), margin = 1)

print('Proportion table by column')
prop.table((xtabs(~term + loan_status, data = loan)), margin = 2)
```


## 9.  The summary all the variables using one command.
```{r}
summary(loan)
```













