---
title: "R_5_visuals_ggplot"
author: "Mike Montes"
date: "2024-01-19"
output: pdf_document
---

```{r setup, include = TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1  Read the dataset in and call it 'loan.' 
```{r warning = FALSE}
library(readr)
loan <- read_csv("C:/Users/mmsax/Downloads/loan.csv")
head(loan, 5)
colnames(loan)
str(loan)
```

# 2  Plot histogram and density of loan_amnt using basic.
## Using extra commands to learn them
```{r}
?plot
?hist

### Histogram
hist(loan$loan_amnt,
     breaks = 'Sturges',
     main = 'Distribution of Loan Amount',
     xlab = 'Amount in $', 
     ylab = 'Count',
     xlim = c(0, 40000),
     ylim = c(0, 1500), 
     col = 'green', 
     border = 'blue',
     labels = TRUE)

### Density
d <- density(loan$loan_amnt)
plot(d,
     col = 'red', 
     lwd = 2)

### Histogram with density overlaid
hist(loan$loan_amnt,
     breaks = 'Sturges',
     main = 'Distribution of Loan Amount',
     prob = TRUE,
     xlab = 'Amount in $', 
     ylab = 'Count',
     xlim = c(0, 40000),
     col = 'green', 
     border = 'blue')
# not sure why I can't have y-limit here
lines(density(loan$loan_amnt),
      lwd = 2,
      col = 'red')
```

# 3  Histogram and density of loan_amnt with vertical line for mean using ggplot2.
```{r warning = FALSE}
?linetype

library(ggplot2)

ggplot(data = loan,
       aes(x = loan_amnt)) +
  geom_histogram(aes(y = ..density..), 
                 color = 'black', 
                 fill = 'yellow') +
  geom_density(alpha = 0.2, 
               fill = 'orange') + 
  geom_vline(aes(xintercept = mean(loan_amnt)),
             color = 'red',
             linetype = 3, 
             size = 1.5) +
  ggtitle('Distribution of Loan Amount') + 
  xlab('Amount in $') +
  ylab('Count')

# I tried to label mean line but couldn't.
#ggplot(df, aes(x=x, y=y)) +
#geom_point() +
 # geom_vline(xintercept=10) +
  #annotate("text", x=9, y=20, label="Some text", angle=90, size=15, color="blue")
```

# 4  The scatter plot of loan_amnt (y-axis) vs. annual_inc (x-axis) + the trend line using basic graphics.
```{r}
?plot

plot(loan$annual_inc, loan$loan_amnt,
     main = 'Scatterplot of Annual Income vs. Loan Amount',
     xlab = 'Annual Income in $',
     ylab = "Loan Amount in $",
     xlim = c(0, 2000000),
     ylim = c(0, 35000),
     pch = 20,
     cex = 0.6)
abline(lm(loan_amnt ~ annual_inc,
          data = loan),
       lwd = 4, 
       col = 'blue3')

# X-axis zoomed 4x
plot(loan$annual_inc, loan$loan_amnt,
     main = 'Scatterplot of Annual Income vs. Loan Amount',
     xlab = 'Annual Income in $',
     ylab = "Loan Amount in $",
     xlim = c(0, 500000),
     ylim = c(0, 35000),
     pch = 20,
     col = 'darkslategrey',
     cex = 0.6)
abline(lm(loan_amnt ~ annual_inc,
          data = loan),
       lwd = 4, 
       col = 'blue3')
```

# 5  Scatter plot of loan_amnt vs.annual_inc with trend line using ggplot2.
```{r}
ggplot(data = loan,
       aes(x = annual_inc, y = loan_amnt)) +
  geom_point() + 
  geom_smooth()

# come back and add to parameters and make a new graph removing the outliers
```

# 6  Barplot of term and grade on the same barplot using basic.
```{r}
# determine if term and grade are numericals
str(loan) 
loan$term <- as.factor(loan$term)
loan$grade <- as.factor((loan$grade))

# determine number of levels to create colors list
levels(loan$term)
levels(loan$grade)

# Grade on x-axis
freq_table_term_grade <- xtabs(~ term + grade,
                               data = loan)
barplot(freq_table_term_grade,
        main = 'Cross Table of Term & Grade',
        legend = rownames(freq_table_term_grade),
        col = c('blue', 'green'),
        ylim = c(0, 3500))

barplot(freq_table_term_grade,
        main = 'Cross Table of Term & Grade',
        legend = rownames(freq_table_term_grade),
        col = c('blue', 'green'), 
        beside = TRUE,
        ylim = c(0, 3000))

# Term on x-axis
freq_table_term_grade_2 <- xtabs(~ grade + term,
                               data = loan)

barplot(freq_table_term_grade_2,
        main = 'Cross Table of Term & Grade',
        legend = rownames(freq_table_term_grade),
        col = c('blue', 'green', 'red', 'pink', 'yellow', 'purple', 'orange'),
        ylim = c(0, 7000))

barplot(freq_table_term_grade_2,
        main = 'Cross Table of Term & Grade',
        legend = rownames(freq_table_term_grade),
        col = c('blue', 'green', 'red', 'pink', 'yellow', 'purple', 'orange'), 
        beside = TRUE,
        ylim = c(0, 3000))
```

# 7  Barplot of term and grade on the same barplot using ggplot2
```{r}
# By term
ggplot(data = loan,
       aes(x = term, y = ..count..)) +
  geom_bar(aes(fill = grade))

ggplot(data = loan,
       aes(x = term, y = ..count..)) +
  geom_bar(aes(fill = grade),
           position = 'dodge')

# By grade
ggplot(data = loan,
       aes(x = grade, y = ..count..)) +
  geom_bar(aes(fill = term))

ggplot(data = loan,
       aes(x = grade, y = ..count..)) +
  geom_bar(aes(fill = term),
           position = 'dodge')
```

# 8  Boxplot loan_amnt vs.term and save as 'loanterm.jpg' using basic graphics.
```{r}
jpeg("C:\\Users\\mmsax\\Desktop\\_MU\\GDSCI_502_R\\Week05\\loanterm.jpg")

boxplot(loan_amnt ~ term, 
        data = loan,
        notch = TRUE,
        col = c('blue'),
        main = 'Loan Amount by Term',
        xlab = 'Term',
        ylim = c(0, 35000))

dev.off()
```

# 9  Boxplot loan_amnt vs.term and save as 'loanterm.jpg' using ggplot2. 
# State differences between loan amount with respect to term.
```{r}
ggplot(data = loan, 
       aes(x = term, 
           y = loan_amnt)) +
  geom_boxplot(aes(col = term), 
               notch = TRUE) +
  ggtitle('Loan Amount vs. Term') +
  xlab('Term Length in Months') +
  ylab('Loan Amount in $') 

ggsave("C:\\Users\\mmsax\\OneDrive\\Desktop\\_MU\\_GDSCI_502_R\\Week05\\loanterm2.jpg", width = 20, height = 15, units = "cm")

print('There is a significant diffence between the average as well as the inner two quartiles of loan amounts between loan length terms of 36 and 60 months.')
```